---
title: "NCAA Women's Basketball Tournament"
author: "Diogo M. Camacho"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidymodels)
library(caret)

```

## Get the data
```{r}
tournament_data <- readr::read_csv(file = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-10-06/tournament.csv")
```


## Exploratory analyses

```{r}
tournament_data <- tournament_data %>%
  dplyr::filter(., !is.na(seed), !is.na(conf_percent))
```

### some plots
```{r year-conference-seed-average}

tournament_data %>%
  dplyr::select(., conference, seed) %>%
  group_by(conference) %>%
  dplyr::count(ts = seed == 1) %>%
  dplyr::filter(ts, n > 1) %>% 
  ggplot(aes(x = forcats::fct_reorder(conference, n), y = n, fill = conference)) + 
  geom_col(color = "black", alpha = 0.8) + 
  coord_flip() +
  labs(title = "Number of #1 seeds per conference", x = NULL, y = NULL) + 
  theme_bw() +
  theme(title = element_text(size = 12, face = "bold"),
        legend.title = element_text(size = 10, face = "plain"))

```


```{r seed-to-reg_percent-wins}
tournament_data %>%
  dplyr::select(., conference, seed, reg_percent) %>%
  group_by(conference) %>%
  ggplot(aes(x = seed, y = reg_percent, color = conference)) + 
  geom_point(alpha = 0.5) +
  facet_wrap(~ conference) +
  theme_bw() +
  theme(title = element_text(size = 12, face = "bold"),
        legend.title = element_text(size = 10, face = "plain"), 
        legend.position = "none")

```

```{r seed-tourn-percent}
tournament_data %>%
  dplyr::select(., conference, seed, full_percent) %>%
  group_by(conference) %>%
  ggplot(aes(x = seed, y = full_percent, color = conference)) + 
  geom_point(alpha = 0.5) +
  facet_wrap(~ conference) +
  theme_bw() +
  theme(title = element_text(size = 12, face = "bold"),
        legend.title = element_text(size = 10, face = "plain"), 
        legend.position = "none")

```


## Prediction of points per team

Let's predict tournament finish given regular season percentage and seed placement. I'll do a regression model based on the generalized linear model with elastic net regularization.

```{r data_split}
data_split <- rsample::initial_split(tournament_data, prop = 0.75)
train_data <- rsample::training(data_split)
test_data <- rsample::testing(data_split)

train_data <- train_data %>%
  dplyr::mutate(tourney_finish = as.factor(tourney_finish))

test_data <- test_data %>%
  dplyr::mutate(tourney_finish = as.factor(tourney_finish))
```


```{r model}
glm_model <- parsnip::linear_reg(mode = "regression", mixture = 0) %>%
  set_engine("glmnet")
```

```{r fit_model}
glm_fit <- glm_model %>%
  parsnip::fit(conf_percent ~ reg_percent + seed, data = train_data)
```

```{r}
glm_preds <- predict(glm_fit, new_data = test_data, penalty = 0.5)
```

```{r compare_predictions}

test_data %>%
  tibble::add_column(glm_pred = glm_preds$.pred) %>%
  ggplot(aes(x = conf_percent, y = glm_pred)) +
  geom_point(alpha = 0.5)

```

## Classification model
Doing a random forest to do multi-class classification on expected tournament finishing. (similar to Julia Silge's datasaurus analyses)

```{r}
# tourney_folds <- train_data %>% 
#   mutate(tourney_finish = factor(tourney_finish)) %>% 
#   bootstraps(times = 50)
```


```{r}
rf_model <- parsnip::rand_forest(trees = 1000, min_n = 5) %>%
  set_mode("classification") %>%
  set_engine("ranger")
```

```{r}
# rf_workflow <- workflows::workflow() %>%
#   add_model(rf_model) %>%
#   add_formula(tourney_finish ~ reg_percent + seed)
```


```{r}
rf_fit <- rf_model %>%
  parsnip::fit(tourney_finish ~ reg_percent + seed, data = train_data)

# doParallel::registerDoParallel()
# rf_fit <- tune::fit_resamples(rf_workflow, tourney_folds, control = control_resamples(save_pred = TRUE))
```

```{r}
rf_preds <- predict(rf_fit, new_data = test_data)
res <- tibble(original = test_data$tourney_finish, prediction = rf_preds$.pred_class)
accuracy <- length(which(res$original == res$prediction)) / nrow(res)

# rf_fit %>% collect_predictions() %>% group_by(id) %>% roc_curve(., .pred_1st:.pred_class) %>% autoplot()

```

